### 2016-11-23 local-capability-machine.tex
More adjustments to the malloc specification

\begin{specification}[Malloc v.2]
  $c_\malloc$ satisfies the specification for malloc iff
  \begin{align*}
    &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
    &\qquad \quad \forall W \future [i \mapsto \iota_{\malloc,0}] \land \\
    &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
    &\qquad \qquad \qquad W' = [i \mapsto \iota_{\malloc}']\land W' \futurewk W \land \\
    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
    &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
    &\qquad \qquad \qquad \size = e'-b' \land\\
  \end{align*}

  Additionally, we require that for every $W' \future [i\mapsto \iota_{\malloc,0}]$, $\heapSat[\hs]{n}{W'}$ implies that $\dom(\hs)$ contains $[n',+\infty)$ for some $n'$.
\end{specification}

@@

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison). If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\item[cclear $r$] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction. In a real system it would probably be implemented as a vector operation which allows modification of continous segments of memory rather fast.}
\item[rclear $\bar{r}$] Moves 0 to all the registers in the list $\bar{r}$.
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

>>>>

\begin{specification}[Malloc v.2]
  $c_\malloc$ satisfies the specification for malloc iff
\[  
  \left\{
  \begin{aligned}
    &\forall \Phi \in \ExecConfs \ldotp \forall m_{\var{footprint}} \in \Heaps\ldotp \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \iota_{\malloc,0} \ldotp \\
    &\qquad \quad \forall W \futurewk [i \mapsto \iota_{\malloc,0}] \land \\
    &\qquad \quad \memheap = m_{\var{footprint}} \uplus \hsframe \land\\
    &\qquad \quad \heapSat[m_{\var{footprint}}]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = c_\malloc \\ %and this needs to change with ABI.
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists m_{\var{footprint}}' \in \Heaps\ldotp \exists \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=m_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
    &\qquad \qquad \qquad W'(i) = \iota_{\malloc}' \land W' \futurewk W \land \\
    &\qquad \qquad \qquad \heapSat[m_{\var{footprint}}']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
    &\qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{(p,b,e,a)}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
    &\qquad \qquad \qquad \size = e'-b' \land\\
  \end{aligned}\right.
\]
where $\iota_{\malloc,0}$ is the region that initially governs malloc 
\end{specification}
In the specfication above $\iota_\malloc'$ is a future region of the initial region that governs malloc.

@@

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.

\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$ (if one of them is an integer, then use that in the comparison). If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\item[cclear $r$] Stores 0 to all the memory cells the capability $r$ governs.\footnote{This may in some cases seem like an unreasonable slow instruction. In a real system it would probably be implemented as a vector operation which allows modification of continous segments of memory rather fast.}
\item[rclear $\bar{r}$] Moves 0 to all the registers in the list $\bar{r}$.
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
